### 2015-07-13 narrative/placement.tex
edits for clarity

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle can be viewed as consisting of four key phases. In the first phase data 
is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). This data has 
to be appropriately initially placed (and possibly partitioned) within the heterogeneous 
multi-tier storage to optimize application I/O overheads.
%In our proposed heterogenous multi-tier storage platform, the system must
%decide where to place the output data at this time, and how to partition and
%reorganize the data to optimize this initial placement. The key optimization
%metrics during this phase is the perceived I/O overhead for the application.

In the second phase of the data life cycle, the data is managed by the system to 
meet application requirements and constraints. This includes migration, 
reorganization and reduction of the data, etc., to optimize resource
utilization, interference with running applications, and availability and
resiliency objectives. 
%Data can be moved into archival storage (tape or cold storage) during this phase. 
%
%In the second phase of the data life cycle the system takes over the
%management of the data and performs the necessary steps to meet service
%level objectives (SLOs). These steps include migration, reorganization, and
%reductions for the data. During this phase the system optimizes for resource
%utilization, interference with running applications, and availability and
%resiliency objectives. Data can be moved into archival storage (tape or cold
%storage) during this phase.

In the third phase the data is consumed for knowledge discovery (for example
read into analytics to produce new scientific insights). It must be noted
that the knowledge discovery process does not remove the data from the
system - the data remains available until it is purged. During this phase
the storage system optimizes for read performance and data availability.

In the final phase the data is purged from the system. The
decision to purge may be made by the system based on operational metrics from
system administrators or by the user. This phase optimizes the space available in the
system. % to prevent out of space concerns for applications.

@@

% Our research will ultimately address the following questions:\\
% 1) Can we do scalable migration without making the discovery process
%     unbounded;\\
% 2) What are the parameters and input for migration;
% 3)  Can we use the additional application level knowledge of data to purge
%     portions of the data from storage without making the data too much less
%     valuable?;\\
% 3)  With so many objects and no centralized directory how do we know what
%     is actually in the storage to be purged?;\\
% 4)  How do purge in a way that doesn't require a central authority,
%     doesn't interfere with other I/O and isn't bottlenecked by some sort of
%     global consistency?

\subsubsection{Data Placement and Movement}
\label{sec:init-plac-data}
When an application outputs or accesses data the storage and middleware layers needs to
decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
This placement decision can have a significant impact on data management 
throughout the life time of the data. For example, our past work on data 
staging~\cite{tongipdps15,qiansc15,docan2012dataspaces,abbasi2010datastager} 
on HPC systems with multi-level memory structures has shown that different 
output techniques targeting different layers of the storage hierarchy can have a 
significant impact on the overhead observed by the application for I/O operations.

@@

A key requirement is application-driven runtime mechanisms for dynamically managing 
data placement across the layers of the distributed storage hierarchy throughout the 
data lifecycle, coordinating data movement and data sharing between the components 
of the application workflow, with the overarching goal of maximizing the relative utility 
to the application as well as the system while reducing access costs. 
%
As noted before, the complexities of heterogeneous multi-level storage structures requires 
adaptive placement policies are required to be implemented to optimally utilize storage 
resources vertically (across deep memory hierarchies) and horizontally (across nodes 
within a memory level) while accommodating dynamic application requirements and transient 
system states. 
%
Our approach is to increase knowledge about the data and its use within the 
application, and leverage this knowledge to drive data placement and overall management.

@@

\paragraph{Soliciting application {\em hints}:}
As described in \S\ref{sec:data-refactor} one of the main components of our
proposed storage system is the ability to reorganize, refactor and reduce data 
as it is generated, and to reorganize and possibly regenerate the data as it is accessed. 
We carry this principle into the placement and movement of data by allowing
applications to define hints and policies that guide \textbf{what} data is
placed \textbf{where}. We will explore the use of application hints in two
distinct areas. First, we will study the challenges and trade-offs of either
augmenting the I/O interface with hints, or allowing the addition of an
external specification that defines the use case. Our experience with
developing modern I/O interfaces has shown that both techniques 
have value~\cite{tongipdps15,qiansc15}, and we will investigate the set of hints that are embedded in the
application code vs. those that are described within an non-compiled
specification. Second, we will study how hints can guide data
placement as data is handed off from application to storage (during a
write) and from storage to application (during the read). In both cases we
will study what minimal set of annotations and hints can allow the storage
system to minimize data movement and optimize the resources consumed by I/O.

@@

\paragraph{ Runtime tracking and estimation:}
Autonomic data placement and/or movement will leverage the information gathered 
at runtime together with predictive analysis requirements. This information may include 
runtime spatial-temporal data access patterns, and/or physical network topology
information, which will be tracked at runtime, and then be used to estimate and predict 
the data access behaviors of application workflow at runtime. 
A key research aspect here will be to anticipate accesses, for example using learning techniques or using
application knowledge, and use this information to prefetch data objects.
For example, we will identify and catalog key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate placement and/or movement actions. Our initial 
work explore these ideas is presented in~\cite{qiansc15,choi2012mining}.

>>>>

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle can be viewed as consisting of four key phases. In the first phase, data 
is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). This data has 
to be appropriately initially placed (and possibly partitioned) within the heterogeneous 
multi-tier storage to optimize application I/O overheads.
%In our proposed heterogenous multi-tier storage platform, the system must
%decide where to place the output data at this time, and how to partition and
%reorganize the data to optimize this initial placement. The key optimization
%metrics during this phase is the perceived I/O overhead for the application.

In the second phase of the data life cycle, the data is managed by the system to 
meet application requirements and constraints. This includes migration, 
reorganization and reduction of the data to optimize resource
utilization, interference with running applications, and availability and
resiliency objectives. 
%Data can be moved into archival storage (tape or cold storage) during this phase. 
%
%In the second phase of the data life cycle the system takes over the
%management of the data and performs the necessary steps to meet service
%level objectives (SLOs). These steps include migration, reorganization, and
%reductions for the data. During this phase the system optimizes for resource
%utilization, interference with running applications, and availability and
%resiliency objectives. Data can be moved into archival storage (tape or cold
%storage) during this phase.

In the third phase, the data is consumed for knowledge discovery. For example
read into analytics to produce new scientific insights. It must be noted
that the knowledge discovery process does not remove the data from the
system - the data remains available until it is purged. During this phase
the storage system optimizes for read performance and data availability.

In the final phase, the data is purged from the system. The
decision to purge may be made by the system based on operational metrics from
system administrators or by the user. This phase optimizes the space available in the
system. % to prevent out of space concerns for applications.

@@

% Our research will ultimately address the following questions:\\
% (1) Can we do scalable migration without making the discovery process
%     unbounded;\\
% (2) What are the parameters and input for migration;
% (3)  Can we use the additional application level knowledge of data to purge
%     portions of the data from storage without making the data too much less
%     valuable?;\\
% (4)  With so many objects and no centralized directory how do we know what
%     is actually in the storage to be purged?;\\
% (5)  How do purge in a way that doesn't require a central authority,
%     doesn't interfere with other I/O and isn't bottlenecked by some sort of
%     global consistency?

\subsubsection{Data Placement and Movement}
\label{sec:init-plac-data}
When an application outputs or accesses data, the storage and middleware layers needs to
decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
This placement decision can have a significant impact on data management 
throughout the life time of the data. For example, our past work on data 
staging~\cite{tongipdps15,qiansc15,docan2012dataspaces,abbasi2010datastager} 
on HPC systems with multi-level memory structures has shown that different 
output techniques targeting different layers of the storage hierarchy can have a 
significant impact on the overhead observed by the application for I/O operations.

@@

A key requirement is application-driven runtime mechanisms for dynamically managing 
data placement across the layers of the distributed storage hierarchy throughout the 
data lifecycle, coordinating data movement and data sharing between the components 
of the application workflow, with the overarching goal of maximizing the relative utility 
to the application as well as the system while reducing access costs. 
%
As noted before, the complexities of heterogeneous multi-level storage structures requires 
adaptive placement policies are required to be implemented to optimally utilize storage 
resources vertically (across deep memory hierarchies) and horizontally (across nodes 
within a memory level) while accommodating dynamic application requirements and transient 
system states. 
%
Our approach is to increase knowledge about the data and its use within the 
application and leverage this knowledge to drive data placement and overall management.

@@

\paragraph{Soliciting application {\em hints}:}
As described in \S\ref{sec:data-refactor}, one of the main components of our
proposed storage system is the ability to reorganize, refactor, and reduce data 
as it is generated and to reorganize and possibly regenerate the data as it is accessed. 
We carry this principle into the placement and movement of data by allowing
applications to define hints and policies that guide \textbf{what} data is
placed \textbf{where}. We will explore the use of application hints in two
distinct areas. First, we will study the challenges and trade-offs of either
augmenting the I/O interface with hints or allowing the addition of an
external specification that defines the use case. Our experience with
developing modern I/O interfaces has shown that both techniques 
have value~\cite{tongipdps15,qiansc15} and we will investigate the set of hints that are embedded in the
application code vs. those that are described within an non-compiled
specification. Second, we will study how hints can guide data
placement as data is handed off from application to storage (during a
write) and from storage to application (during the read). In both cases we
will study what minimal set of annotations and hints can allow the storage
system to minimize data movement and optimize the resources consumed by I/O.

@@

\paragraph{ Runtime tracking and estimation:}
Autonomic data placement and/or movement will leverage the information gathered 
at runtime together with predictive analysis requirements. This information may include 
runtime spatial-temporal data access patterns and/or physical network topology
information, which will be tracked at runtime and then be used to estimate and predict 
the data access behaviors of application workflow at runtime. 
A key research aspect here will be to anticipate accesses, for example, using learning techniques or using
application knowledge and use this information to prefetch data objects.
For example, we will identify and catalog key phases and access patterns
in the target workflows and then use it to characterize patterns at runtime
and trigger appropriate placement and/or movement actions. Our initial 
work explore these ideas is presented in~\cite{qiansc15,choi2012mining}.
